### 2021-10-26 paper/pldi22/formal.tex
Syntatic views in formalization

Since $\ires$ is a specification language for JavaScript, it treats JavaScript
ASTs $\treeset$ as its values. For presentation brevity, we represent a JavaScript
AST as follows:
\[
  \tree ::= \str \mid \ty_k \langle \tree^* \rangle
\]
A string $\str$ denotes a leaf node with a terminal symbol, and $\ty_k \langle
\tree^* \rangle$ denotes a non-leaf node for $k$-th alternative in the syntactic
production of nonterminal symbol $\ty$ with its subtrees $\tree^*$.  The notation
$\ty_k.\eval$ denotes an $\ires$ function for the evaluation of $k$-th
alternative in $\ty$, and it takes the AST itself and its nonterminal subtrees
as arguments. For each integer $k$, the notation $\tree[k]$ denotes $k$-th
subtree of $\tree$.  For example, Figure~\ref{fig:coalesce-prod} shows a
syntactic production for coalesce expressions.  Consider the following coalesce
expression:
\begin{lstlisting}[style=JS]
                    42 ?? true
\end{lstlisting}
Then, the following AST is produced as its parsing result:
\begin{figure}[H]
  \includegraphics[width=.8\columnwidth]{img/ast-example.png}
\end{figure}
\noindent Moreover, its evaluation function $\nterm{CoalesceExpression}_0.\eval$
takes three subtrees as arguments annotated by $\tree_0$, $\tree_1$, and
$\tree_2$ in the AST.

@@

        \aenv \join \aenv' &=& \lambda \varx \in \varset. \; \aenv(\varx) \join
        \aenv(\varx)\\
      \end{array}
    \]
  \item \underline{Abstract Values}: $\avalset = (\valset \times \exprset)
    \uplus \{ \top \}$
    \[
      \begin{array}{lcl}
        \multicolumn{3}{l}{\valgamma: \avalset \rightarrow \powerset{\valset}}\\
        \valgamma(\aval) &=& \left\{
          \begin{array}{ll}
            \{ \val \} & \aval = (\val, \expr) \\
            \valset & \aval = \top\\
          \end{array}
        \right.\\

@@

\subsubsection{Expressions} Before defining a transformation of expressions, we
first define abstract evaluation of expressions as follows:
\[
  \framebox{$\asem{\expr}: \aenvset \rightarrow \avalset$}
\]
\begin{itemize}
  \item \underline{Primive Values}:
    \[
      \asem{\pval}(\aenv) = \langle \pval, \pval \rangle
    \]
  \item \underline{Operations}:
    \[
      \asem{\op \kwrl \expr_1, \cdots, \expr_n \kwrr}(\aenv) = \aval
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}l@{~}l}
        \asem{\expr_k}(\aenv) &=& \aval_k \; \forall 1 \leq k \leq n
        &\wedge\\
        \aval &=& \left\{
          \begin{array}{ll}
            \top & \text{if} \; \exists k. \; \aval_k = \top\\
            \langle \val, \expr' \rangle & \text{if} \; \forall k. \; \aval_k =
            \langle \val_k, \expr'_k \rangle\\
          \end{array}
        \right. &\wedge\\
        \val &=& \op(\val_1, \cdots, \val_n) &\wedge\\
        \expr' &=& \op(\expr'_1, \cdots, \expr'_n)\\
      \end{array}
    \]
  \item \underline{Variable Lookups}:
    \[
      \asem{\varx}(\aenv) = \aenv(\varx)
    \]
  \item \underline{Field Lookups}:
    \[
      \asem{\expr_0 \kwsl \expr_1 \kwsr}(\aenv) = \aval
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}l@{~}l}
        \asem{\expr_0}(\aenv) &=& \aval_0 &\wedge\\
        \asem{\expr_1}(\aenv) &=& \aval_1 &\wedge\\
        \aval &=& \left\{
          \begin{array}{ll}
            \langle \tree[k], \expr'_0 \kwsl \expr'_1 \kwsr \rangle & \text{if}
            \; \aval_0 = \langle \tree, \expr'_0 \rangle \wedge \aval_1 =
            \langle k, \expr'_1 \rangle\\
            \top & \text{otherwise}\\
          \end{array}
        \right.\\
      \end{array}
    \]
\end{itemize}
Then, we define the transformation of expressions as follows:
\[
  \framebox{$\transforme: \exprset \times \aenvset \rightarrow \exprset \times
  \avalset$}
\]
\[
  \transforme(\expr, \aenv) = (\expr'', \aval)
\]
where
\[
  \begin{array}{l@{~}c@{~}ll}
    \aval &=& \asem{\expr}(\aenv) &\wedge\\
    \expr'' &=& \left\{
      \begin{array}{ll}
        \pval & \text{if} \; \aval = \langle \pval, \_ \rangle\\
        \expr' & \text{if} \; \aval = \langle \_, \expr' \rangle\\
        \expr & \text{otherwise}\\
      \end{array}
    \right.\\
  \end{array}
\]

\subsubsection{Instructions} Now, we define a transformation of instructions
using the transformation of expressions as follows:
\[
  \framebox{$\transformi: \instset \times \aenvset \rightarrow \instset \times
  \aenvset$}
\]
\begin{itemize}
  \item \underline{Variable Assignments}:
    \[
      \transformi(\varx = \expr, \aenv) =
      (\varx = \expr', \aenv[\varx \mapsto \aval])
    \]
    where
    \[
      (\expr', \aval) = \transforme(\expr, \aenv)
    \]

@@

\subsubsection{Fixpoint of Abstract Transfer Function}

@@

This section explains how to utilize the constant propagation specialization for
the JavaScript semantics specialization with a user-defined syntactic view.  We
first formally define syntactic views and their operations, then we explain how
to decide a restriction of the constant propagation specialization based on a
given syntactic view.

\subsubsection{Syntatic Views}

@@

\subsubsection{Specialization with Syntatic Views}

>>>>

Since $\ires$ is a specification language for JavaScript, it treats JavaScript
ASTs $\treeset$ as its values. For presentation brevity, we represent a JavaScript
AST as follows:
\[
  \tree ::= \str \mid \ty_k \langle \tree^* \rangle
\]
A string $\str$ denotes a leaf node with a terminal symbol, and $\ty_k \langle
\tree^* \rangle$ denotes a non-leaf node for $k$-th alternative in the syntactic
production of nonterminal symbol $\ty$ with its subtrees $\tree^*$.  The notation
$\ty_k.\eval$ denotes an $\ires$ function for the evaluation of $k$-th
alternative in $\ty$, and it takes the AST itself and its nonterminal subtrees
as arguments. For each integer $k$, the notation $\tree[k]$ denotes $k$-th
subtree of $\tree$. If an AST $\tree$ is a subtree of another AST $\tree'$, we
represent it as $\tree \subtree \tree'$.  For example,
Figure~\ref{fig:coalesce-prod} shows a syntactic production for coalesce
expressions.  Consider the following coalesce expression:
\begin{lstlisting}[style=JS]
                    42 ?? true
\end{lstlisting}
Then, the following AST is produced as its parsing result:
\begin{figure}[H]
  \includegraphics[width=.8\columnwidth]{img/ast-example.png}
\end{figure}
\noindent Its evaluation function $\nterm{CoalesceExpression}_0.\eval$ takes
three subtrees as arguments annotated by $\tree_0$, $\tree_1$, and $\tree_2$ in
the figure. Note that $\tree_1$ and $\tree_2$ are subtrees of $\tree_0$ (i.e.,
$\tree_1 \subtree \tree_0$ and $\tree_2 \subtree \tree_0$).

@@

        \aenv \join \aenv' &=& \lambda \varx \in \varset. \; \aenv(\varx) \join
        \aenv(\varx)\\
      \end{array}
    \]
  \item \underline{Abstract Values}: $\avalset = (\valset \times \exprset)
    \uplus \{ \top \}$
    \[
      \begin{array}{lcl}
        \multicolumn{3}{l}{\valgamma: \avalset \rightarrow \powerset{\valset}}\\
        \valgamma(\aval) &=& \left\{
          \begin{array}{ll}
            \{ \val \} & \text{if} \; \aval = (\val, \expr) \\
            \valset & \text{if} \; \aval = \top\\
          \end{array}
        \right.\\

@@

\subsubsection{Transformation of Expressions} Before defining a transformation of
expressions, we first define abstract evaluation of expressions as follows:
\[
  \framebox{$\asem{\expr}: \aenvset \rightarrow \avalset$}
\]
\begin{itemize}
  \item \underline{Primive Values}:
    \[
      \asem{\pval}(\aenv) = \langle \pval, \pval \rangle
    \]
  \item \underline{Operations}:
    \[
      \asem{\op \kwrl \expr_1, \cdots, \expr_n \kwrr}(\aenv) = \aval
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}l@{~}l}
        \asem{\expr_k}(\aenv) &=& \aval_k \; \forall 1 \leq k \leq n
        &\wedge\\
        \aval &=& \left\{
          \begin{array}{ll}
            \top & \text{if} \; \exists k. \; \aval_k = \top\\
            \langle \val, \expr' \rangle & \text{if} \; \forall k. \; \aval_k =
            \langle \val_k, \expr'_k \rangle\\
          \end{array}
        \right. &\wedge\\
        \val &=& \op(\val_1, \cdots, \val_n) &\wedge\\
        \expr' &=& \op(\expr'_1, \cdots, \expr'_n)\\
      \end{array}
    \]
  \item \underline{Variable Lookups}:
    \[
      \asem{\varx}(\aenv) = \aenv(\varx)
    \]
  \item \underline{Field Lookups}:
    \[
      \asem{\expr_0 \kwsl \expr_1 \kwsr}(\aenv) = \aval
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}l@{~}l}
        \asem{\expr_0}(\aenv) &=& \aval_0 &\wedge\\
        \asem{\expr_1}(\aenv) &=& \aval_1 &\wedge\\
        \aval &=& \left\{
          \begin{array}{ll}
            \langle \tree[k], \expr'_0 \kwsl \expr'_1 \kwsr \rangle & \text{if}
            \; \aval_0 = \langle \tree, \expr'_0 \rangle \wedge \aval_1 =
            \langle k, \expr'_1 \rangle\\
            \top & \text{otherwise}\\
          \end{array}
        \right.\\
      \end{array}
    \]
\end{itemize}
Then, we define the transformation of expressions as follows:
\[
  \framebox{$\transforme: \exprset \times \aenvset \rightarrow \exprset \times
  \avalset$}
\]
\[
  \transforme(\expr, \aenv) = (\expr'', \aval)
\]
where
\[
  \begin{array}{l@{~}c@{~}ll}
    \aval &=& \asem{\expr}(\aenv) &\wedge\\
    \expr'' &=& \left\{
      \begin{array}{ll}
        \pval & \text{if} \; \aval = \langle \pval, \_ \rangle\\
        \expr' & \text{if} \; \aval = \langle \_, \expr' \rangle\\
        \expr & \text{otherwise}\\
      \end{array}
    \right.\\
  \end{array}
\]

\subsubsection{Transformation of Instructions} Now, we define a transformation
of instructions using the transformation of expressions as follows:
\[
  \framebox{$\transformi: \instset \times \aenvset \rightarrow \instset \times
  \aenvset$}
\]
\begin{itemize}
  \item \underline{Variable Assignments}:
    \[
      \transformi(\varx = \expr, \aenv) =
      (\varx = \expr', \aenv[\varx \mapsto \aval])
    \]
    where
    \[
      (\expr', \aval) = \transforme(\expr, \aenv)
    \]

@@

\subsubsection{Fixpoint of Abstract Transfer Function} Finally, we define an
abstract transfer function and its fixpoint to represent abstract semantics of a
given function and abstract values:
\[
  \begin{array}{ll}
    \text{Abstract Transfer Function} & \atransfer: \aelemset \rightarrow
    \aelemset\\
\end{array}
\]

@@

This section explains how to utilize the partial evaluation for the JavaScript
semantics specialization with a user-defined syntactic view.  We first formally
define syntactic views and their operations, then we explain how to decide a
restriction of the constant propagation specialization based on a given
syntactic view.

\subsubsection{Syntatic Views} A syntatic view $\atree \in \atreeset$ is an
augmented JavaScript ASTs with abstract nodes $\top$ as follows:
\[
  \atree ::= \str \mid \ty_k \langle \atree^* \rangle \mid \top
\]
\[
  \begin{array}{lcl}
    \multicolumn{3}{l}{\treegamma: \atreeset \rightarrow \powerset{\treeset}}\\
    \treegamma(\atree) &=& \left\{
      \begin{array}{ll}
        \{ \str \} &
        \text{if} \; \atree = \str \\

        A & \text{if} \; \atree = \ty_k
        \langle \atree_1, \cdots, \atree_n \rangle \\

        \treeset & \atree = \top\\
      \end{array}
    \right.\\
    A &=& \{ \ty_k \langle \tree_1, \cdots, \tree_n \rangle \mid \forall 1 \leq
    j \leq n.  \; \tree_j \in \treegamma(\atree_j) \}
  \end{array}
\]

\subsubsection{JavaScript Program Filtering}

\[
  \atree \subtree \tree' \iff \exists \tree \in \treegamma{\atree}. \; \tree
  \subtree \tree'
\]

@@

\subsubsection{Semantics Specialization}
